EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-06-01 03_edipack2.tex
v1 + few corrections Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\subsubsection{Impurity Hamiltonian}
The impurity Hamiltonian is split into a quadratic and an 
interacting part,
\begin{equation}\label{Himp}
  \hat{H}_\mathrm{imp}  = \hat{H}^0_\mathrm{imp} + \hat{H}^\mathrm{int}_\mathrm{imp}.
\end{equation}
The first term has the form
\begin{equation}\label{H0imp}
  \hat{H}^0_\mathrm{imp}  =
  \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^\dagger_{\a\sigma}d_{\b\sigma'}
\end{equation}
and represents the non-interacting part, with $d_{\a\sigma}$ ($d^\dagger_{\a\sigma}$) being the annihilation (creation)
operators for impurity electrons in 
orbital $\alpha=1,\dots,N_\a$, where $N_\a$ is the number of orbitals,
and spin $\sigma=\up,\dw$.
The internal structure of this quadratic part is captured by the 
matrix $h^{0}_{\a\b\sigma\sigma'}$, which can include 
orbital-dependent on-site energies or amplitudes, spin-orbit coupling or other single-particle terms.
In \NAME the matrix $\hat{h}_0$ is set by the user passing an complex array through the function {\tt ed\_set\_Hloc}.

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction \cite{Georges2013ARCMP}:
\begin{equation}\label{Hint}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^\dagger_{\a\uparrow}d_{\a\downarrow}d^\dagger_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^\dagger_{\a\uparrow}d^\dagger_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow},\\
\end{split}
\end{equation}
where we introduced the occupation number operator $n_{\alpha\sigma}=d^\dagger_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling \cite{Werner2008PRL,Werner2009PRB,Mravlje2011PRL,Medici2011PRL,Medici2011PRB,Georges2013ARCMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$ \cite{Georges2013ARCMP}. Different choices, preserving part
of the combined symmetry group, can be made for other numbers of orbitals \cite{Georges2013ARCMP}.

@@

\paragraph{{\tt bath\_type}={\bf hybrid}.}
The hybrid topology is the simplest generalization of the bath pa\-ra\-me\-tri\-za\-tion which captures the effects of locally
coupled impurity orbitals. However, this setup typically requires a more challenging optimization process, especially
with few available bath levels (see \secu{sSecFit}). 
The bath is formed by $N_\mathrm{bath}$ sites coupled to all the
impurity orbitals, so that there are $N_\mathrm{b}\equiv  N_\mathrm{bath}$ bath levels and   $N_\mathrm{s}=N_\mathrm{bath}+N_\a$. 
The parametrization of the bath Hamiltonian is 
diagonal in both orbital and spin space, i.e.
$$
h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'}.
$$
For {\tt ed\_mode}={\bf superc} additional anomalous components 
$\Delta_\a^p\delta_{\a\b}\delta_{\s\bar{\s}}$ diagonal in the orbital basis are included.
%

>>>>

\subsubsection{Impurity Hamiltonian}
The impurity Hamiltonian is split into a quadratic and an 
interacting part,
\begin{equation}\label{Himp}
  \hat{H}_\mathrm{imp}  = \hat{H}^0_\mathrm{imp} + \hat{H}^\mathrm{int}_\mathrm{imp}.
\end{equation}
The first term has the form
\begin{equation}\label{H0imp}
  \hat{H}^0_\mathrm{imp}  =
  \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^\dagger_{\a\sigma}d_{\b\sigma'}
\end{equation}
and represents the non-interacting part, with $d_{\a\sigma}$ ($d^\dagger_{\a\sigma}$) being the annihilation (creation)
operators for impurity electrons in 
orbital $\alpha=1,\dots,N_\a$, where $N_\a$ is the number of orbitals,
and spin $\sigma=\up,\dw$.
The internal structure of this quadratic part is captured by the 
matrix $h^{0}_{\a\b\sigma\sigma'}$, which can include 
orbital-dependent on-site energies or amplitudes, spin-orbit coupling or other single-particle terms.
In \NAME the matrix $\hat{h}_0$ is set by the user passing a complex array through the function {\tt ed\_set\_Hloc}.

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can, in principle, contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction \cite{Georges2013ARCMP}:
\begin{equation}\label{Hint}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^\dagger_{\a\uparrow}d_{\a\downarrow}d^\dagger_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^\dagger_{\a\uparrow}d^\dagger_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow},\\
\end{split}
\end{equation}
where we introduced the occupation number operator $n_{\alpha\sigma}=d^\dagger_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling \cite{Werner2008PRL,Werner2009PRB,Mravlje2011PRL,Medici2011PRL,Medici2011PRB,Georges2013ARCMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$ \cite{Georges2013ARCMP}. Alternative choices that preserve part
of the combined symmetry group can be made for other numbers of orbitals \cite{Georges2013ARCMP}.

@@

\paragraph{{\tt bath\_type}={\bf hybrid}.}
The hybrid topology is the simplest generalization of the bath pa\-ra\-me\-tri\-za\-tion which captures the effects of locally
coupled impurity orbitals. However, this setup typically requires a more challenging optimization process, especially
with few available bath levels (see \secu{sSecFit}). 
The bath is formed by $N_\mathrm{bath}$ sites coupled to all the
impurity orbitals, so that there are $N_\mathrm{b}\equiv  N_\mathrm{bath}$ bath levels and   $N_\mathrm{s}=N_\mathrm{bath}+N_\a$. 
The parametrization of the bath Hamiltonian is 
diagonal in both orbital and spin spaces, i.e.
$$
h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'}.
$$
For {\tt ed\_mode}={\bf superc} additional anomalous components 
$\Delta_\a^p\delta_{\a\b}\delta_{\s\bar{\s}}$ diagonal in the orbital basis are included.
%
